A GRASP algorithm for the concrete delivery problem.

### 2023-08-15 Article/article.tex
Save article

A driver typically loads RMC at his assigned batch plant but may be required to drive to and load at other plants if needed. The batch plant produces concrete on demand using recipes specific to each order. This means that a truck can only haul RMC for one order, even if there is spare capacity.  After unloading the RMC,  driver $k$ takes $Ct_k$ minutes to clean the concrete mixer before proceeding. %When assigning drivers to deliveries, the company prioritizes employees with the highest seniority.

@@

Let the binary variable $x^{k}_{ij}$ be $1$ if driver $k$ travels from node $i$ to $j$. Binary variable $y_o$ is $1$ when order $o$ is completely served; $v_i$ and $w_i$ are the start and end of the loading (unloading) operation at node $i \in \mathcal{L}\cup \mathcal{D}$. Binary variable $u_{ij}$ is $1$ if node $j$ is served just after $i$, the service being either an unloading or a loading operation. % Binary variable $\sigma_{ib}$ is equal to 1 if the orders of customer $i$ are loaded by plant $b$. 
Variable $q^k_j$ is the quantity to be loaded towards $j$ with vehicle $k$. Let $w^1_k$ be a continuous variable indicating the difference between the driver's work time and the minimum number of hours to be worked in a day, and $w^2_k$ indicating the difference between the driver's work time and the normal work time. Let $g_i$ be the time between due date and first service start for customer $i$.
% Binary variable $u_{o_{1},o_{2}}$ equals $1$ if order $o_{2}$ is served directly after order $o_{1}$. Binary variables $z_{o,d}$ ($z_{i,d}$) indicates whether or not delivery $d \in  \mathcal{D}_o$ ($d \in  \mathcal{D}_i$) is supposed to be executed. $w_i$ is the end of the loading (unloading) operation at node $i \in L\cup D$.

The objective function minimizes the travel cost, the penalty costs incurred when customer demands are not fully met, the lateness of each customer's first delivery, the cost of drivers working less than the minimum hours, and the total overtime cost of drivers working beyond their scheduled hours. The mathematical model of this variant of the CDP is as follows:

%The objective function (\ref{mod:obj}) is designed to minimize multiple factors that impact the overall cost and efficiency of the concrete delivery process. These factors include: The total traveled distance, which represents the fuel consumption and wear and tear on the delivery vehicles. The penalty costs $\beta_1$, which are incurred when customers' demands are not fully satisfied.The lateness of each customer's first delivery, which can impact customer satisfaction and lead to additional penalty costs.The number of drivers who work under the minimum working time, which can lead to increased labor costs and decreased productivity.The total overtime cost, which is incurred when drivers work beyond their scheduled hours.All these factors are taken into consideration in the objective function to minimize the cost and maximize the satisfaction of the customers.

@@

    \label{mod:obj}   \min & \smashoperator{ \sum_{ \substack{(i,j,k) \in A} }} {t_{ij}x^{k}_{ij}} + \beta_1\smashoperator{\sum_{ o \in \mathcal{O}}}{  \left(1-y_o\right)} \hspace{-0mm}  + \beta_2 \smashoperator{ \sum_{i \in \mathcal{C} } } { g_i  } +  \sum_{k \in K}{\beta_3 w^{1}_k } + \beta_4 { w^{2}_k } &                                                                                                  \\

@@

    \label{mod:c4}         & v_j \geq  w_i + Ct_{k} + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                           & \hspace{-20mm} \forall i \in \mathcal{D}  , j \in \delta^{+}(i), k \in K                         \\

@@

\section{Computational results}
\label{comp_exp}

In this section, we present and discuss the results of the computational experiments of the GRASP heuristic described in section~\ref{grasp_method}. The GRASP algorithm is coded in C++. We run our experiments on a dataset extracted from data provided by our industrial partner.

\subsection{Generation of instances}

The dataset used in this section was obtained from our partner and contains records representing delivery operations from up to 8 production centers over a period of 50 days. We generated 50 instances, with each instance corresponding to a daily operation. The instances include informations such as each driver capacity, associated depot, and shift start time. For each customer, we have the due time, total demand, and the number of orders (type of concrete) received. For each order, we specified its demand and associated production plant. Finally, we include loading capacities for each production center. 
The name of an instance has the format $C\_k\_n\_o\_z$. $k$ is the number of available drivers, $n$ the number of customers, $o$ the total number of daily orders, and $z$ the number of plants.

The dataset is subdivised in a small, medium and large sets, according to the number of drivers and total demands served. We summarized both sets in Table~\ref*{}
or 
The dataset is subdivised in a small, medium and large sets, according to the number of orders served. We summarized both sets in Table~\ref*{}

% This dataset allows us to evaluate the effectiveness and performance of our GRASP algorithm in optimizing delivery operations in a real-world setting.

\begin{itemize}
    \item   8 production centers located in greater Montreal
    \item   42 days of delivery operations from Fall 2020
    \item   From 7 to 139 orders each day
    \item   Up to 137 drivers
    \item   2 types of trucks of capacity 8 and 12 m$^3$
\end{itemize}

\subsubsection{ Results for the generated instances}

- resultats avec les instances où toute la demande est satisfaite. Montrer la distance parcourue, le temps de première livraison, overtime, underwork, temps d'execution.

- resultats avec les instances (8 au total) où uniquement la flotte de Unibéton a éét utilisée. (utiliser les mêmes pour la comparaison avec les données réelles )
distance, demande servie, first, underwork, overtime.
- resultats avec les autres instances
- distance, demande servie, first delivery.

% df = df[df['Name'].isin([20201010,20201021,20201107,20201017,20201031,20201109,20201026,20201121])]

\begin{table}
    \resizebox*{\textwidth}{!}{

\begin{tabular}{lrrrrrrr}
    \toprule
    {} &  satisfiedCost &  driverUsed &  travelCost &  firstDeliveryCost &  underWorkCost &  overTimeCost &  runtime \\
    Name     &                &             &             &                    &                &               &          \\
    \midrule
    20201121 &          226.5 &          13 &     1243.55 &               0.51 &           0.00 &          0.00 &     5.26 \\
    20201010 &          267.0 &          13 &      745.10 &               0.00 &           0.00 &          0.00 &     5.32 \\
    20201107 &          333.5 &          18 &      777.25 &               0.00 &         200.16 &          0.00 &     5.19 \\
    20201031 &          375.0 &          15 &     1017.64 &               0.00 &         216.06 &         28.10 &     5.42 \\
    20201114 &          388.0 &          19 &     1743.64 &               0.00 &         149.48 &          0.00 &     5.42 \\
    20201128 &          776.0 &          31 &     2278.18 &               0.00 &           4.19 &        438.14 &     5.16 \\
    20201024 &          937.5 &          35 &     4403.57 &               0.00 &          58.30 &       1005.24 &     5.72 \\
    20201021 &         1160.5 &          73 &     6107.22 &               0.00 &        2535.40 &        185.02 &     5.28 \\
    20201013 &         1565.0 &         102 &     9214.74 &              19.53 &        2955.35 &         95.74 &     5.32 \\
    20201026 &         1839.5 &         116 &    13585.60 &              44.44 &        1826.70 &        227.56 &     5.41 \\
    20201007 &         2512.5 &         127 &    15166.00 &               5.09 &         877.68 &        574.76 &     5.67 \\
    \bottomrule
    \end{tabular}
    }
\end{table}

\subsubsection{ Results for the instances of \cite{kinable2014concrete}}

\cite{kinable2014concrete} provided a benchmark dataset for their variant of the $CDP$. Although their primary objective is to maximize the total load delivered each day, without considering loading operations at a plant and ensuring deliveries within specified time windows, we can still use this dataset to analyze and compare the performance of our algorithm under different conditions. Table \ref*{tab:cdp_instances} shows the performance of the GRASP heuristic along with the other methods (CP, MIP, SD-heuristic, and Hybrid) used in \cite{kinable2014concrete}.  For each method, the table shows the average execution time in milliseconds and the corresponding gap between the obtained solution and the upper bound provided by \cite{kinable2014concrete}.  Overall, the GRASP algorithm shows promise in providing near-optimal solutions in a relatively short time for this benchmark of the CDP. It outperformed all methods except CP, achieving an average distance of 5.7\% (13.8\%) with a runtime of 13,186 (574,983) milliseconds for Set A (Set B). The SD-heuristic appears to be much faster, but GRASP provides a better balance between solution quality and computation time. It should also be noted that the exact MIP, CP, and Hybrid are initialized with the results of the SD-heuristic.

\begin{table}[htbp]
    \centering
    \caption{Performance of the GRASP heuristic on the CDP instances}
    \label{tab:cdp_instances}
    \resizebox{1\textwidth}{!}{
    \begin{tabular}{@{}cccccccccccc@{}}
    \toprule
     &  & \multicolumn{2}{c}{CP} & \multicolumn{2}{c}{MIP} & \multicolumn{2}{c}{SD-heuristic} & \multicolumn{2}{c}{Hybrid} & \multicolumn{2}{c}{GRASP} \\ \cmidrule(l){3-10} \cmidrule(l){11-12} 
     &  & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) \\ \cmidrule(l){3-10} \cmidrule(l){11-12} 
    \multirow{2}{*}{Set A} & Avg & 4.2 & 197,012 & 7.3 & 13,405,798 & 9.1 & 23 & 7.0 & 100,765 & 5.7 & 13,186 \\ \cmidrule(l){2-10} \cmidrule(l){11-12} 
     & Opt & \multicolumn{2}{c}{40/64} & \multicolumn{2}{c}{37/64} & \multicolumn{2}{c}{30/64} & \multicolumn{2}{c}{35/64} & \multicolumn{2}{c}{35/64} \\ \cmidrule(l){3-10} \cmidrule(l){11-12} 
     \multirow{2}{*}{Set B} & Avg & 12.1 & 357,313 & - & - & 16.3 & 1,331 & - & - & 13.8 & 574,983\\ \cmidrule(l){2-10} \cmidrule(l){11-12} 
     & Opt & \multicolumn{2}{c}{55/128} & \multicolumn{2}{c}{-} & \multicolumn{2}{c}{40/128} & \multicolumn{2}{c}{-} & \multicolumn{2}{c}{44/128} \\ \bottomrule
    \end{tabular}
    }
    \end{table}

@@

- rouler les instances sans shift des livreurs. comparer la distance parcourue, et demande livrée.
- rouler sans les livreurs externes. regarder la demande servie par jour.

@@

>>>>

A driver typically loads RMC at his assigned batch plant but may be required to drive to and load at other plants if needed. The batch plant produces concrete on demand using recipes specific to each order. This means that a truck can only haul RMC for one order, even if there is spare capacity.  After unloading the RMC,  driver $k$ takes $\rho$ minutes to clean the concrete mixer before proceeding. %When assigning drivers to deliveries, the company prioritizes employees with the highest seniority.

@@

Let the binary variable $x^{k}_{ij}$ be $1$ if driver $k$ travels from node $i$ to $j$. Binary variable $y_o$ is $1$ when order $o$ is completely served; $v_i$ and $w_i$ are the start and end of the loading (unloading) operation at node $i \in \mathcal{L}\cup \mathcal{D}$. Binary variable $u_{ij}$ is $1$ if node $j$ is served just after $i$, the service being either an unloading or a loading operation. % Binary variable $\sigma_{ib}$ is equal to 1 if the orders of customer $i$ are loaded by plant $b$. 
Variable $q^k_j$ is the quantity to be loaded towards $j$ with vehicle $k$. Let $w^1_k$ be a continuous variable indicating the difference between the driver's work time and the minimum number of hours to be worked in a day, and $w^2_k$ indicating the difference between the driver's work time and the normal work time. Let $g_i$ be the time between due date and first service start for customer $i$.

The objective function minimizes total travel cost ($TC$), penalties associated with unfulfilled orders ($PUO$), first delivery delays ($FDD$), driver underutilization costs ($DUC$), and driver overtime costs ($DOC$). Together, these components drive the optimization process to find a solution that efficiently balances travel costs, customer satisfaction, on-time delivery, driver utilization, and scheduling constraints.

% The objective function minimizes the total travel cost, the penalty costs incurred when customer demands are not fully met, the lateness of each customer's first delivery, the cost of drivers working less than the minimum hours, and the total overtime cost of drivers working beyond their scheduled hours. The mathematical model of this variant of the CDP is as follows:

@@

    \label{mod:obj}   \min & \smashoperator{ \sum_{ \substack{(i,j,k) \in A} }} {t_{ij}x^{k}_{ij}} + \beta_1\smashoperator{\sum_{ o \in \mathcal{O}}}{  \left(1-y_o\right)} \hspace{-0mm}  + \beta_2 \smashoperator{ \sum_{i \in \mathcal{C} } } { g_i  } +  \sum_{k \in K}\left({\beta_3 w^{1}_k  + \beta_4  w^{2}_k}\right)  &                         \\

@@

    \label{mod:c4}         & v_j \geq  w_i + \rho + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                           & \hspace{-20mm} \forall i \in \mathcal{D}  , j \in \delta^{+}(i), k \in K                         \\

@@

\section{Computational experiments}
\label{comp_exp}

In this section, we present and discuss the results of computational experiments conducted using the GRASP heuristic described in section~\ref{grasp_method}. The implementation of the GRASP algorithm is coded in C++. We run our experiments on the benchmark data used in \cite{kinable2014concrete}, and on a dataset extracted from delivery operations records provided by our industry partner.

\subsection{Generation of instances}
\newcommand{\nbInstance}{36}
The dataset used in this section was obtained from our partner and contains records representing delivery operations from up to 8 production centers over a period of $\nbInstance$ days. We have created $\nbInstance$ instances, where each instance corresponds to a daily operation. An instance contains information such as each driver's capacity, associated batch plant, and shift start time. For each customer, we have the due time, the total demand, and the number of orders (type of concrete) received. For each order, we have its demand and associated production plant. Finally, we have the loading capacity for each plant. The name of an instance has the format $C\_k\_n\_o\_z$, where $k$ is the number of available drivers, $n$ is the number of customers, $o$ is the total number of daily orders, and $z$ is the number of plants. The dataset is divided into small, medium, and large sets according to the total number of daily orders. We have summarized the dataset in the Table~\ref{tab:instances}.

\begin{table}[htpb]
    \centering
    \caption{Instances summary}
    \label{tab:instances}
    \resizebox*{0.8\textwidth}{!}{
    \begin{tabular}{lcccccc}
        \toprule
        {} &  \# &         Demand  &   \#Order & \#Client &  \#Driver & \#Depot \\
          &       &         (m$^3$)        &         &         &          &        \\
        \midrule
        Small  &     8 &    226.5-937.5 &     7-14 &    4-11 &    13-35 &    1-3 \\
        Medium &    14 &  1,160.5-2,971.0 &    43-98 &   40-89 &   76-137 &    6-8 \\
        Large  &    14 &  3,078.5-3,953.5 &  107-136 &  92-127 &  129-150 &    8 \\
        \bottomrule
        \end{tabular}  
    }
\end{table}

\subsubsection{Results for the instances of \cite{kinable2014concrete}}

\cite{kinable2014concrete} provided a benchmark dataset for their variant of the $CDP$. Although their primary objective is to maximize the total load delivered each day, without considering loading operations at a plant and ensuring deliveries within specified time windows, we still use this dataset to analyze and compare the performance of our algorithm under different conditions. Table \ref*{tab:cdp_instances} shows the performance of the GRASP heuristic along with the other methods (CP, MIP, SD-heuristic, and Hybrid) used in \cite{kinable2014concrete}.  For each method, the table shows the average execution time in milliseconds and the corresponding gap between the obtained solution and the upper bound provided by \cite{kinable2014concrete}.

\begin{table}[htbp]
    \centering
    \caption{Performance of the GRASP heuristic on the CDP benchmark instances}
    \label{tab:cdp_instances}
    \resizebox{1\textwidth}{!}{
    \begin{tabular}{@{}cccccccccccc@{}}
    \toprule
     &  & \multicolumn{2}{c}{CP} & \multicolumn{2}{c}{MIP} & \multicolumn{2}{c}{SD-heuristic} & \multicolumn{2}{c}{Hybrid} & \multicolumn{2}{c}{GRASP} \\ \cmidrule(l){3-10} \cmidrule(l){11-12} 
     &  & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) \\ \cmidrule(l){3-10} \cmidrule(l){11-12} 
    \multirow{2}{*}{Set A} & Avg & 4.2 & 197,012 & 7.3 & 13,405,798 & 9.1 & 23 & 7.0 & 100,765 & 5.7 & 13,186 \\ \cmidrule(l){2-10} \cmidrule(l){11-12} 
     & Opt & \multicolumn{2}{c}{40/64} & \multicolumn{2}{c}{37/64} & \multicolumn{2}{c}{30/64} & \multicolumn{2}{c}{35/64} & \multicolumn{2}{c}{35/64} \\ \cmidrule(l){3-10} \cmidrule(l){11-12} 
     \multirow{2}{*}{Set B} & Avg & 12.1 & 357,313 & - & - & 16.3 & 1,331 & - & - & 13.8 & 574,983\\ \cmidrule(l){2-10} \cmidrule(l){11-12} 
     & Opt & \multicolumn{2}{c}{55/128} & \multicolumn{2}{c}{-} & \multicolumn{2}{c}{40/128} & \multicolumn{2}{c}{-} & \multicolumn{2}{c}{44/128} \\ \bottomrule
    \end{tabular}
    }
\end{table}

Overall, the GRASP algorithm shows promise in providing near-optimal solutions in a relatively short time for this benchmark of the CDP, making it a suitable candidate for further analysis and real-world application. It outperformed all methods except CP, achieving an average distance of 5.7\% (13.8\%) with a runtime of 13,186 (574,983) milliseconds for Set A (Set B). The SD-heuristic appears to be much faster, but GRASP provides a better balance between solution quality and computation time. It should also be noted that the exact MIP, CP, and Hybrid are initialized with the results of the SD-heuristic.

\subsubsection{Results for the generated instances}

The computational results of our instances solved with the GRASP heuristic are summarized in the tables below. The stopping criterion is set to 20,000 iterations. The parameters used in our experiments are listed in Table~\ref*{tab:problem_parameters}.

\begin{table}[htpb]
    \centering
    \caption{Parameters of the real-world instances}
    \label{tab:problem_parameters}
%    \resizebox*{\textwidth}{!}{
    \begin{tabularx}{\textwidth}{Xr}
        \toprule
        \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Value} \\ \midrule
        Unloading duration ($\tau_c^u$) & 2 min$/m^3$ \\
        Cleaning duration ($\rho$) & 10 min \\
        Adjustment duration ($\alpha$) & 10 min \\
        Max travel time ($\Delta$) & 120 min \\
        Min working time ($M_T$) & 180 min \\
        Max working time ($N_T$) & 480 min \\
        Max overtime duration ($O_T$) & 120 min \\
        Max delay between two consecutive deliveries of the same order ($\gamma^1$) & 20 min \\
        Max delay between two consecutive deliveries of different orders ($\gamma^2)$ & 25 min \\
    %    $\beta_1$ & 10000  \\
    %    $\beta_2$ & 10000  \\
    %    $\beta_3$ & 1000  \\
    %    $\beta_4$ & 100  \\
        \bottomrule
    \end{tabularx}
%    }
\end{table}

Table \ref*{tab:runtime} shows the percentage of concrete delivered by varying the execution times of the algorithm. For small instances, all requests are delivered in less than five minutes, while for medium and large instances, over 99\% of requests are delivered.  For these instances, it takes more than ten minutes to ensure complete delivery of all orders.

\begin{table}[htbp]
    \centering
    \caption{RMC delivered within different runtimes}
    \label{tab:runtime}
    \resizebox*{0.5\textwidth}{!}{
        \begin{tabular}{llcccc}
            \toprule
            &  &      \multicolumn{4}{c}{Runtime (min)} \\
            \cmidrule{3-6}
            & &      5   & 10  &  30  & 60  \\
            %  & {} &         &         &        &        \\
            \midrule
            \multirow{1}{*}{Small} & \% Load &  100.00 &  100.00 &  100.0 &  100.0 \\
                %   & \# instances &    8 &    8 &    8 &    8 \\
            \cmidrule{2-6}
                  \multirow{1}{*}{Medium} & \% Load &   99.86 &   99.86 &  100.0 &  100.0 \\
                %   & \# instances &   13 &   13 &   14 &   14 \\
                  \cmidrule{2-6}
                  \multirow{1}{*}{Large} & \% Load &   99.79 &   99.93 &  100.0 &  100.0 \\
                %   & \# instances &   12 &   13 &   14 &   14 \\
            \bottomrule
            \end{tabular} 
    }

\end{table}

Table \ref*{tab:results_grasp_ls} shows the detailed results for each instance. The runtime is set to 3600 seconds. The table reports the requirements of each instance, the components of the objective function, the number of drivers used, and the execution time in seconds.
%  the travel cost ($TC$), the penalties associated with unfulfilled orders ($PUO$), the sum of the first delivery delays ($FDD$) for each customer, the sum of the driver underutilization costs ($DUC$) and driver overtime costs ($DOC$).

All daily orders are served with our algorithm, using less drivers than the available fleet. For the small size instances, all first deliveries are timely, except for one instance. The sum of the driver underutilization costs ($DUC$) is high for medium and large instances. This indicates underutilization of the scheduled fleet. Some drivers are not scheduled at all or work very little. The sum of the driver overtime costs ($DOC$) is also high for these instances. This can be explained by the fact that the algorithm prefers to put certain drivers on overtime rather than use a driver who has to travel to another plant for a delivery, which would increase the travel cost ($TC$).

$FDD$ is the sum of the delays of the first deliveries. To get an idea of the actuel delay per customer, we report $mFDD$, which is the maximum delay for all first deliveries. This tells us that we have a maximum delay that varies between 0 and 12 minutes for small instances, 1 and 60 minutes for medium instances, and 50 to 60 minutes for large instances.

\begin{table}[htbp]
    \caption{Results with all instances}
    \label{tab:results_grasp_ls}
    \resizebox*{\textwidth}{!}{
        \begin{tabular}{llccccccccc}
        \toprule
        &   &  Demand &  $TC$ & $PUO$&  $FDD$ & $mFDD$ &  $DUC$ &  $DOC$ &  \#Drivers &  Time \\
        & \multicolumn{1}{c}{Instance} &  (m$^3$)  &(min)  &  & (min) &(min) &(min)    & (min) & & (s)\\
        \midrule
            \multirow{9}{*}{Small} & C\_13\_11\_12\_1 &          226.5 &     1243.55 &  0.0 &               0.00 &              0.00 &          66.01 &          0.00 &          13 &  3022.06 \\
                  & C\_13\_5\_8\_1 &          267.0 &      745.10 &  0.0 &               0.00 &              0.00 &           0.00 &          0.00 &          13 &  3027.39 \\
                  & C\_18\_6\_11\_2 &          333.5 &      753.26 &  0.0 &               0.00 &              0.00 &         653.78 &          0.00 &          17 &  3004.51 \\
                  & C\_15\_4\_7\_2 &          375.0 &     1013.11 &  0.0 &               0.00 &              0.00 &          56.79 &        244.61 &          15 &  3009.54 \\
                  & C\_19\_7\_8\_2 &          388.0 &     1714.74 &  0.0 &               0.00 &              0.00 &          97.56 &        283.83 &          19 &  3020.89 \\
                  & C\_29\_10\_14\_3 &          613.5 &     3724.29 &  0.0 &              12.41 &             12.41 &         246.04 &        454.51 &          28 &  3035.23 \\
                  & C\_31\_8\_11\_3 &          776.0 &     2297.20 &  0.0 &               0.00 &              0.00 &         627.37 &        539.96 &          30 &  3033.85 \\
                  & C\_35\_9\_10\_3 &          937.5 &     4077.20 &  0.0 &               0.00 &              0.00 &         720.00 &        866.98 &          31 &  3020.75 \\
                  \cmidrule{2-11}
                  \multirow{14}{*}{Medium} & C\_76\_40\_43\_6 &         1160.5 &     5715.53 &  0.0 &               1.00 &              1.00 &        3566.49 &        771.23 &          69 &  3021.92 \\
                  & C\_104\_42\_47\_8 &         1565.0 &     9114.62 &  0.0 &              23.00 &             16.24 &        4180.95 &        671.94 &          94 &  3047.99 \\
                  & C\_94\_63\_70\_7 &         1746.5 &    13888.90 &  0.0 &              19.86 &             12.40 &        1086.60 &       1477.76 &          94 &  3025.72 \\
                  & C\_116\_67\_78\_8 &         1839.5 &    13946.30 &  0.0 &              44.68 &              9.98 &        1578.90 &        639.28 &         116 &  3007.88 \\
                  & C\_116\_71\_82\_8 &         2060.0 &    12461.20 &  0.0 &              68.38 &             20.06 &        3119.66 &       1478.63 &         107 &  3028.24 \\
                  & C\_117\_57\_70\_6 &         2327.5 &    15066.80 &  0.0 &              83.79 &             16.44 &        1522.11 &       1892.64 &         117 &  3020.70 \\
                  & C\_137\_79\_83\_7 &         2425.0 &    18811.40 &  0.0 &             252.94 &             41.65 &        2193.63 &       1963.05 &         136 &  3025.83 \\
                  & C\_127\_66\_80\_8 &         2512.5 &    15177.80 &  0.0 &               2.00 &              2.00 &        1874.83 &        965.65 &         125 &  3007.70 \\
                  & C\_128\_68\_74\_7 &         2595.0 &    19133.50 &  0.0 &             251.77 &             54.40 &        1523.19 &       2924.28 &         128 &  3053.96 \\
                  & C\_136\_85\_97\_8 &         2673.0 &    16832.10 &  0.0 &             154.97 &             50.21 &        3507.40 &       1958.58 &         132 &  3028.15 \\
                  & C\_128\_78\_85\_8 &         2685.5 &    16927.30 &  0.0 &             202.26 &             50.12 &        1492.50 &       2794.21 &         128 &  3030.32 \\
                  & C\_131\_77\_85\_8 &         2893.5 &    17572.10 &  0.0 &             187.42 &             49.24 &        1803.22 &       1306.94 &         130 &  3040.74 \\
                  & C\_137\_89\_97\_7 &         2939.5 &    21934.10 &  0.0 &             697.18 &             59.10 &        1220.15 &       3038.33 &         137 &  3025.45 \\
                  & C\_133\_84\_98\_7 &         2971.0 &    22399.30 &  0.0 &             867.12 &             59.44 &         550.77 &       2887.51 &         133 &  3137.11 \\
                  \cmidrule{2-11}
                  \multirow{14}{*}{Large} & C\_132\_98\_109\_8 &         3078.5 &    23102.40 &  0.0 &             193.01 &             49.44 &         450.57 &       3931.83 &         132 &  3018.05 \\
                  & C\_129\_101\_119\_8 &         3229.0 &    18978.90 &  0.0 &             467.94 &             59.47 &        1084.78 &       2334.16 &         129 &  3179.81 \\
                  & C\_141\_114\_136\_8 &         3350.5 &    21760.00 &  0.0 &            1131.39 &             59.58 &        1284.13 &       2360.19 &         141 &  3056.56 \\
                  & C\_140\_114\_132\_8 &         3401.5 &    23259.00 &  0.0 &            1092.82 &             57.99 &         634.53 &       2705.26 &         140 &  3046.75 \\
                  & C\_143\_101\_123\_8 &         3437.5 &    24560.70 &  0.0 &             451.36 &             56.61 &        1663.43 &       4359.64 &         142 &  3039.02 \\
                  & C\_137\_112\_129\_8 &         3471.0 &    21526.10 &  0.0 &             252.62 &             58.09 &         624.87 &       3360.64 &         137 &  3018.76 \\
                  & C\_142\_114\_129\_8 &         3499.5 &    21991.50 &  0.0 &             898.10 &             55.22 &        1750.60 &       4943.03 &         142 &  3011.54 \\
                  & C\_149\_98\_122\_8 &         3513.0 &    20696.60 &  0.0 &             839.56 &             52.60 &        2465.87 &       2892.55 &         149 &  3072.58 \\
                  & C\_139\_98\_108\_8 &         3541.0 &    20550.10 &  0.0 &             566.69 &             57.49 &        1438.11 &       2910.97 &         139 &  3011.84 \\
                  & C\_144\_108\_122\_8 &         3670.5 &    23401.60 &  0.0 &             668.15 &             59.77 &        1438.60 &       3569.62 &         143 &  3056.51 \\
                  & C\_142\_92\_107\_8 &         3684.5 &    22700.20 &  0.0 &             147.42 &             48.31 &         899.82 &       3702.01 &         142 &  3019.52 \\
                  & C\_138\_114\_136\_8 &         3739.5 &    24670.70 &  0.0 &             750.04 &             55.88 &         518.63 &       5156.41 &         137 &  3021.26 \\
                  & C\_150\_127\_136\_8 &         3946.5 &    25822.20 &  0.0 &             742.75 &             57.10 &         899.92 &       5101.58 &         150 &  3060.79 \\
                  & C\_148\_112\_131\_8 &         3953.5 &    25455.30 &  0.0 &             868.81 &             49.15 &         500.68 &       3775.57 &         148 &  3005.96 \\
            \bottomrule
            \end{tabular}
\end{table}

We tried to reduce the maximum delay of the first delivery but found that some orders couldn't be fulfilled because of driver shifts and overtime constraints. We decided to remove these constraints and rerun our computation. We found that for small instances, the maximum delay dropped to zero, and for medium and large instances, it was reduced by almost 50\%. This suggests that when we input a good work schedule, our algorithm works effectively to ensure that all orders are filled and that delays for first deliveries and travel costs are minimized. In Table~\ref*{tab:lift_shift_constraints}, we can see that the average travel time decreases for all instances except the large ones. That's because the goals of minimizing travel cost and not delaying the first delivery are sometimes in conflict: to ensure that the first delivery is not late, drivers assigned to one plant often have to travel to another. And since all plants are used in large cases, this can increase travel times.

%  Therefore, we can use our algorithm for fleet sizing or as a submodule of a driver scheduling algorithm.

\begin{table}[htbp]
    \centering
    \caption{Comparaison with and without shift and overtime constraints}
    \label{tab:lift_shift_constraints}
    \resizebox*{0.8\textwidth}{!}{
        \begin{tabular}{lcccccccc}
            \toprule
        {} & \multicolumn{4}{c}{Default parameters} &\multicolumn{4}{c}{$M_T=0$, $N_T,O_T=\infty$, $H_k=0$} \\
        \cmidrule(l){2-5} \cmidrule(l){6-9}
        {} &   Load & $TC$ &   $FDD$ &  $mFDD$ &   Load &      $TC$ &   $FDD$ &   $mFDD$ \\
               &($m^3$)  & (min)   &(min)   &(min)  &($m^3$)   &  (min) &(min)  & (min)   \\
            \midrule
            Small  &   489.62 &   1946.06 &    1.55 &       1.55 &   489.62 &   1832.90 &    0.00 &       0.00 \\
            Medium &  2313.86 &  15641.50 &  204.03 &      31.59 &  2313.86 &  14607.66 &   74.76 &      16.56 \\
            Large  &  3536.86 &  22748.24 &  647.90 &      55.48 &  3536.86 &  22293.04 &  210.14 &      29.00 \\
            \bottomrule
            \end{tabular}

\end{table}

The influence of the parameter $\gamma^1$ on the solution quality is demonstrated in Figure~\ref*{fig:gamma1_influence}. Tests were conducted by varying $\gamma^1$ values between 10 and 25 minutes, without any driver shifts or overtime constraints. In small instances, all customers are served regardless of the value of $\gamma^1$. For medium (large) instances, all customers are served with $\lambda^1$ starting at 15 (20) minutes. The delay of the initial delivery decreases as the value of $\gamma^1$ increases. The results indicate that if $\gamma^1$ is high, the plant has the flexibility to delay ongoing deliveries to make room for new ones or make customers wait until the drivers are available. By increasing the maximum delay between two successive deliveries from 10 to 20 minutes, the first delivery delay is reduced by over 50\% for the medium and large size instances. The average travel time remains constant for all instances, regardless of the value of $\gamma^1$.

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{gamma1.pdf} 
    \caption{Influence of the maximum time delay between consecutive deliveries of the same order ($\gamma^1$)}
    \label{fig:gamma1_influence}
\end{figure}

We checked how well the local search part of the GRASP algorithm works by running tests using only constructive heuristics. The results are presented in Table~\ref*{tab:grasp_ls_nols}. For medium and large instances, the use of local search was necessary to service all customers within 60 minutes. And without local search, most objective metrics experienced a significant increase, except for travel costs.

\begin{table}[htbp]
    \centering
    \caption{Influence of the local search step on the  GRASP algorithm}
    \label{tab:grasp_ls_nols}
    \resizebox*{1\textwidth}{!}{
    \begin{tabular}{lccccccccccccc}
        \toprule
           & &\multicolumn{5}{c}{GRASP$\backslash$LS} & \multicolumn{5}{c}{GRASP}   \\
           \cmidrule(l){2-14}
        {size} &Load &&$TC$&   $FDD$ &  $DUC$ &  $DOC$ & Time&&  $TC$ &  $FDD$&  $DUC$ &  $DOC$&Time \\
        {} & (m$^3$)&\%Served&(min)&   (min) &  (min) &  (min) & (min)&\%Served&  (min) &  (min)&  (min) &  (min)& (min) \\
        \cmidrule(l){2-2} \cmidrule(l){3-8} \cmidrule(l){9-14}
        Small  & 3917.0& 100 &       1998.75 &   64.60 &           231.63 &          215.43 &21.24& 100&      1946.06 &    1.55 &    229.40 &          298.74 &50.34\\
        Medium & 32394.0& 99.6&     13144.05 &   2528.33 &  2496.47 & 1086.12 &49.41&100&  13564.08 &   75.77 &   2250.66 &         1347.59&50.6 \\
        Large  & 49516.0& 99.2& 21395.23 & 4885.11 &  1250.14 &  3543.16 & 50.20&100&     21784.95 &  471.62 & 1168.73 &  3640.04&50.74 \\
        \bottomrule
        \end{tabular}
    }  
\end{table}

@@

- resultats avec les instances (8 au total) où uniquement la flotte de Unibéton a éét utilisée.
enlver les contraintes de temps. reduire le gamma1

(utiliser les mêmes pour la comparaison avec les données réelles )
distance, demande servie, first, underwork, overtime.
- resultats avec les autres instances
- distance, demande servie, first delivery.
- rouler les instances sans shift des livreurs. comparer la distance parcourue, et demande livrée.
- rouler sans les livreurs externes. regarder la demande servie par jour.

@@

Il serait intéressant de voir l'effet des paramètres lambda 1 et lambda 2 différents par clients et /ou ordre. Également le temps de service différent par client.

l'algorithme serait intéressant à utiliser dans le cadre d'un

%  Therefore, we can use our algorithm for fleet sizing or as a submodule of a driver scheduling algorithm.
